---
title: "Bivariate gamma mixture of experts models"
author: "Sen Hu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bivariate gamma mixture of experts models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `mvClaim` package provides a flexible modelling framework of mixture of experts (MoE) using bivariate gamma distributions, as introduced in Hu et al. (2019).
It implements the bivariate gamma distribution proposed by Cheriyan (1941) and Ramabhadran (1951), which has not received much attention in the past. 
Depends on the parameterization of the framework, its interpretation, and model parsimony issue, a family of models are implemented including: 

* Bivariate gamma distribution estimation
* Bivariate gamma regression
* Mixture of bivariate gamma clustering
* Mixture of bivariate gamma regressions (i.e. mixture of bivariate gamma clustering with covariates)


In this documents we illustrate two examples of bivariate gamma mixture of experts (MoE) models with two artificually simulated data sets, as well as illustrating some main functions in the `mvClaim` package. 


## Example 1







## Example 2

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
